Sketch of DP - Bayesian Inference

### 2018-07-17 main-sketch.tex
remvoed a sentence not useful

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}\mathrm{x_i}$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
where $\Delta \alpha_i=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha).
\]
%where $+$ denotes the

\section{The Problem and Baseline approach}
We are interested in designing a mechanism for privately releasing
\emph{fully} the posterior distributions derived in section \ref{sec_background},
and not a sample from them.  It's worth noticing that the posterior
distributions are fully characterized by their parameters, and the
family (beta, Dirichlet) they belong to.  Hence, in case of the
Beta-Binomial model we are interested in releasing a private version
of the pair of parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a
private version of $\valpha'=(\valpha + \Delta \valpha)$.  \cite{} and \cite{} have
already attacked this problem and their solution consisted in adding
independent Laplacian noise to the parameters of the posteriors. That
is, in the case of the Beta-Binomial system the value released would
be: $(\alpha + \Delta \tilde\alpha,\beta + n - \Delta \tilde\alpha)$ where $\Delta \tilde\alpha\sim Lap(\Delta \alpha, \frac{2}{\epsilon})$ with truncating method come from algorithm designed by \citet{zhang2016differential}
.  Where $Lap(\mu,\nu)$
denotes a Laplacian random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^{n}$ we say that $\dataobs$ and $\dataobs'$ are adjacent iff
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$.} data observations
$\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system.

\noindent Also, in previous works, the accuracy has been measured again with respect to
$\ell_1$ norm. That is, an lower bound has been given to
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as  defined above.
In this work we will use a metric based on a different norm to compute the sensitivity
and provide guarantees on the accuracy. In particular we will consider a metric
over probability measures and not over the parameters that represent them.
Specifically, we will use the Hellinger distance, which for two beta distributions
$\betad(\alpha_1, \beta_1), \betad(\alpha_2, \beta_2)$ is defined as
\[
  \hellinger()\equiv\sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}.
\]
The same change of metric will be applied to the experimental accuracy guarantees.
\section{Technical Problem Statement}
\label{sec_bayesInfer}
The posterior belief about parameters $\xis$ is inferred from the
group of sensitive data. Because it is usually published in form of a
distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself
can relieve a lot of information about sensitive data. So, the
sensitive information we want to protect here is the posterior belief,
i.e., $Pr(\xis | \dataobs)$, rather than just a sample drew from
it\cite{foulds2016theory}. Then it will be better if we can use a
distance over two distributions to measure the accuracy than a
distance over just parameters or values when we are protecting a
distribution, for instance, Hellinger distance in our paper. When we
measure the accuracy between true posterior and protected posterior
using Hellinger distance, it would be better we can have a protection
mechanism based on Hellinger distance. However, current mechanisms are
mainly based on $l_1$ norm by Laplace mechanism
\cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we
design our exponential mechanism over Hellinger distance to protect
the whole posterior distribution.

>>>>

\section{Bayesian Inference Background}
\label{sec_background}
Given a prior belief $\Pr(\theta)$ on some parameter $\theta$,
and given an observation $\dataobs$,
the posterior distribution on $\theta$ given $\dataobs$ is computed as:
\[
  \Pr(\theta | \dataobs) = \frac{\Pr(\dataobs | \theta) \cdot \Pr(\theta)}{\Pr(\dataobs)}
\]
where the expression  $\Pr(\dataobs | \theta)$ denotes the \emph{likelihood} of $\theta$ when
$\dataobs$ is observed. Since we consider $\dataobs$ to be fixed the likelihood is a function of $\theta$.
For the same reason $\Pr(\dataobs)$ is a constant independent of $\theta$.
Usually in statistics the prior distribution $\Pr(\theta)$ is chosen so that it represents
the initial belief on $\theta$. That is, when no data has been observed. In practice though,
prior distributions and likelihood functions are usually chosen so that the posterior
belongs to the same \emph{family} of distributions. In this case we say that the prior
is conjugate of the likelihood function. Using conjugate priors simplify calculcations and allows
for inference to be performed in a recursive fashion over the data.
In this work we will consider a specific instance of Bayesian inference and one of its generalizations.
Specifically, we will consider the situation where $\theta$ represents the parameter --informally called \emph{bias}--
of a Bernoulli distributed random variable, and its immediate generalization where the parameter $\vtheta$ represents
the the vector of parameters of a categorical distributed random variable.
In the former case the prior distribution over $\theta\in [0,1]$ is going to be a beta distribution, $\betad(\alpha, \beta)$, with parameters
$\alpha,\beta\in\mathbb{R}^{+}$, and with p.d.f:
\[
  \Pr(\theta)\equiv \frac{\theta^{\alpha} (1- \theta)^{\beta}}{\betaf(\alpha,\beta)}
\]
where $\betaf(\cdot,\cdot)$ is the beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ 0/1 values, that is $\dataobs=\ll x_1,\dots x_n \gg, x_i\in\{0,1\}$, and the likelihood function is:
\[
  \Pr(\dataobs | \theta)\equiv \theta^{\Delta \alpha}(1-\theta)^{n - \Delta \alpha}
\]
where $\Delta \alpha = \displaystyle\sum_{i=1}^{n}x_i$.
From this it can easily be derived that the posterior distribution is:
\[
  \Pr(\theta|\dataobs)=\betad(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)
\]
In the latter case the prior distribution over $\vtheta\in [0,1]^{k}$ is given by a Dirichelet distribution, $\dirichlet(\valpha)$, for $k\in\mathbb{N}$,
and $\valpha\in(\mathbf{R^{+}})^{k}$, with p.d.f:
\[
  \Pr(\vtheta)\equiv\frac{1}{\mbetaf(\valpha)}\cdot \displaystyle\prod_{i=1}^{k}{\theta_i^{\alpha_i-1}}
\]
where $\mbetaf(\cdot)$ is the generealised beta function.
The data $\dataobs$ will be a sequence of $n\in\mathbb{N}$ values coming from a universe $\datauni$, such that $\mid\datauni \mid=k$.
The likelihood function will be:
\[
  \Pr(\dataobs|\vtheta)\equiv \displaystyle\prod_{a_i\in\datauni}\theta_{i}^{\Delta \alpha_i},
\]
where $\Delta \alpha_i=\displaystyle\sum_{i=1}^{n}\iverson{x_i=a_i}$, where in $\iverson{\cdot}$ we use
Iverson bracket notation. The posterior distribution over $\vtheta$ turns out to be
\[
  \Pr(\vtheta|\dataobs)=\dirichlet(\valpha+\Delta \valpha).
\]
%where $+$ denotes the

\section{The Problem and Baseline approach}
We are interested in designing a mechanism for privately releasing
\emph{fully} the posterior distributions derived in section \ref{sec_background},
and not a sample from them.  It's worth noticing that the posterior
distributions are fully characterized by their parameters, and the
family (beta, Dirichlet) they belong to.  Hence, in case of the
Beta-Binomial model we are interested in releasing a private version
of the pair of parameters $(\alpha',\beta')=(\alpha + \Delta \alpha,\beta + n - \Delta \alpha)$, and
in the case of the Dirichlet-Multinomial model we are interested in a
private version of $\valpha'=(\valpha + \Delta \valpha)$.  \cite{} and \cite{} have
already attacked this problem and their solution consisted in adding
independent Laplacian noise to the parameters of the posteriors. That
is, in the case of the Beta-Binomial system the value released would
be: $(\alpha + \Delta \tilde\alpha,\beta + n - \Delta \tilde\alpha)$ where $\Delta \tilde\alpha\sim Lap(\Delta \alpha, \frac{2}{\epsilon})$.
%with truncating method come from algorithm designed by \citet{zhang2016differential}
Where $Lap(\mu,\nu)$
denotes a Laplacian random variable with mean $\mu$ and scale $\nu$.
This mechanism is $\epsilon$-differentially private, and the noise is
calibrated w.r.t to a sensitivity of 2 which is derived by using
$\ell_1$ norm over the pair of parameters. Indeed, considering two
adjacent\footnote{Given $\dataobs, \dataobs'\in\{0,1\}^{n}$ we say that $\dataobs$ and $\dataobs'$ are adjacent iff
$\displaystyle \sum_{i}^{n}\iverson{x_i = x'_i }\leq 1$.} data observations
$\dataobs, \dataobs'$, that, from a unique prior, give rise to two posterior
distributions, characterized by the pairs
$(\alpha',\beta')$ and $(\alpha'',\beta'')$ then
$|\alpha'-\alpha''|+|\beta'-\beta''|\leq 2$.
This argument extends similarly to the Dirichelet-Multinomial system.

\noindent Also, in previous works, the accuracy has been measured again with respect to
$\ell_1$ norm. That is, a lower bound has been given to
\[
  \Pr[|\alpha - \tilde\alpha| + |\beta - \tilde\beta |\geq \gamma ]
\]
where $(\alpha, \beta), (\tilde\alpha,\tilde\beta)$ are as  defined above.
In this work we will use a metric based on a different norm to compute the sensitivity
and provide guarantees on the accuracy. In particular we will consider a metric
over probability measures and not over the parameters that represent them.
Specifically, we will use the Hellinger distance, which for two beta distributions
$\boldsymbol{\beta}_1=\betad(\alpha_1, \beta_1), \boldsymbol{\beta}_2=\betad(\alpha_2, \beta_2)$ is defined as
\[
  \hellinger(\boldsymbol{\beta}_1, \boldsymbol{\beta}_2)\equiv
  \sqrt{1 - \frac{\betaf(\frac{\alpha_1 + \alpha_2}{2}, \frac{\beta_1 + \beta_2}{2})}{\sqrt{\betaf(\alpha_1,\beta_1)\betaf(\alpha_2,\beta_2)}}}
\]
The same change of metric will be applied to the experimental accuracy guarantees.
\section{Technical Problem Statement}
\label{sec_bayesInfer}
The posterior belief about parameters $\xis$ is inferred from the
group of sensitive data. Because it is usually published in form of a
distribution, for example, $\betad(\alpha',\beta'), \cdots,$ itself
can relieve a lot of information about sensitive data. So, the
sensitive information we want to protect here is the posterior belief,
i.e., $Pr(\xis | \dataobs)$, rather than just a sample drew from
it\cite{foulds2016theory}. Then it will be better if we can use a
distance over two distributions to measure the accuracy than a
distance over just parameters or values when we are protecting a
distribution, for instance, Hellinger distance in our paper. When we
measure the accuracy between true posterior and protected posterior
using Hellinger distance, it would be better we can have a protection
mechanism based on Hellinger distance. However, current mechanisms are
mainly based on $l_1$ norm by Laplace mechanism
\cite{zhang2016differential} \cite{xiao2012bayesian}. That's why we
design our exponential mechanism over Hellinger distance to protect
the whole posterior distribution.
